Efficient Handling of String-Number Conversion

### 2019-11-23 draft.tex
add example

The correctness of the entire construction of $\underf \dr {\insc}$ is expressed by  Theorem~\ref{theorem:correct}. 
It uses the decoding function $\decode \dr$ parameterized by the domain restriction $\dr$. 
Let $Z$ be the set of integer variables in $\insc$.
The function maps an interpretations $I_e$ over $Z\cup V_\dr\cup\#V_\dr$
to an interpretation over $Z \cup X$, following the domain restriction $\dr$. Informally, it ``decodes" an interpretation of integer variables $Z\cup V_\dr\cup\#V_\dr$ to an interpretation of variables in the string constraint $\insc$. Formally,
we define $\decode  \dr (I_e) \defeq \{I \mid \forall z\in Z: I(z) = I_e(z) \wedge\forall x\in X: \{I(x)\} = \decode {\dr(x)} ((I_e)_{V_{\dr(x)}\cup \#V_{\dr(x)}})\}$.
The condition says that (1) $I$ and $I_e$ are consistent over variables in $Z$ and (2) $(I_e)_{V_{\dr(x)}\cup \#V_{\dr(x)}}$ is a word encoding that $\dr(x)$-encodes $I(x)$.
%We define the output of $\decode  \dr (I_e\cup I_z)$ as a set so that \emph{invalid word encodings}, i.e. those do not corresponds to possible word encodings from PFAs, are mapped to empty sets and valid ones are mapped to singleton sets.
We also define the $\dr$-encoding function as the counterpart of $\dr$-decoding,  
%
namely, for a interpretation $I$ of the string constraint $\insc$,  
we let $\encode \dr (I) =\{I_e\mid \decode \dr (I_e) = I \}$.
We lift $\decode \dr$ and $\encode \dr$ to sets of interpretations in the standard manner.
\begin{theorem}\label{theorem:correct}
$\decode \dr (\semof {\underf \dr {\insc}}) = \semof{\insc}^\dr$ 
\end{theorem}

@@

Formally, the asynchronous product automaton is a tuple $\syncof \pa {\pa'} = (Q\times Q', \syncT, \cvarone_\epsilon \times \cvartwo_\epsilon, (q_i,q_i'),(q_f,q_f'))$, where the transition relation $\syncT$ is the minimal set satisfying the following:
%
\begin{itemize}
\item If $(q_1,v,q_2) \in T$ and $(q_1',v',q_2') \in T'$, then we have $((q_1,q_1'),(v,v'),(q_2,q_2'))\in \syncT$.
\item If $(q_1,v,q_2) \in T$, then for all states $q'\in Q'$, we have $((q_1,q'),(v,\epsilon),(q_2,q'))\in \syncT$.
\item If $(q_1',v',q_2') \in T'$, then for all states $q\in Q$, we have $((q,q_1'),(\epsilon,v'),(q,q_2'))\in \syncT$.
\end{itemize}	
%
The synchronization formula $\syncfof{\pa} {\pa'}$ is extracted from $\syncof \pa {\pa'}$ as follows.
Its first part is the Parikh formula $\parikhfof {\syncof \pa {\pa'}}$ of the product, which encodes all accepting runs of $\syncof \pa {\pa'}$.
%
The second part is a constraint that extracts from a run of $\syncof \pa {\pa'}$ the corresponding runs of $\pa$ and of $\pa'$:
$$ \Psi_{\#} \defeq 
\left(\bigwedge_{v\in\cvarone}\#v = \sum_{x' \in \cvartwo_\epsilon} \#{(v,x')}\right)
\land
\left(\bigwedge_{v'\in\cvartwo} \#v' = \sum_{x \in \cvarone_\epsilon} \#{(x,v')}\right)
$$
Notice that $x,x'$ are either variables or $\epsilon$.
Finally, the third part forces the interpretations of the parametric words accepted by $\pa$ and $\pa'$ to be the same:
$$ \Psi_= \defeq
\bigwedge_{x\in\cvarone_\epsilon,x'\in\cvartwo_\epsilon} \#(x,x')>0 \rightarrow (x=x')
$$
The synchronization formula is then the conjunction 
$$
\syncfof \pa {\pa'} \defeq \parikhfof{\syncof \pa {\pa'}} \land \Psi_{\#} \land \Psi_= \land \paf \land \paf'
$$

>>>>

The correctness of the entire construction of $\underf \dr {\insc}$ is expressed by  Theorem~\ref{theorem:correct}. 
It uses the decoding function $\decode \dr$ parameterized by the domain restriction $\dr$. 
Let $Z$ be the set of integer variables in $\insc$.
The function maps an interpretations $I_e$ over $Z\cup V_\dr\cup\#V_\dr$
to an interpretation over $Z \cup X$, following the domain restriction $\dr$. Informally, it ``decodes" an interpretation of integer variables $Z\cup V_\dr\cup\#V_\dr$ to an interpretation of variables in the string constraint $\insc$. Formally,
we define $\decode  \dr (I_e) \defeq \{I \mid \forall z\in Z: I(z) = I_e(z) \wedge\forall x\in X: \{I(x)\} = \decode {\dr(x)} ((I_e)_{V_{\dr(x)}\cup \#V_{\dr(x)}})\}$.
The condition says that (1) $I$ and $I_e$ are consistent over variables in $Z$ and (2) $(I_e)_{V_{\dr(x)}\cup \#V_{\dr(x)}}$ is a word encoding that $\dr(x)$-encodes $I(x)$.
%We define the output of $\decode  \dr (I_e\cup I_z)$ as a set so that \emph{invalid word encodings}, i.e. those do not corresponds to possible word encodings from PFAs, are mapped to empty sets and valid ones are mapped to singleton sets.
We also define the $\dr$-encoding function as the counterpart of $\dr$-decoding,  
%
namely, for a interpretation $I$ of the string constraint $\insc$,  
we let $\encode \dr (I) =\{I_e\mid \decode \dr (I_e) = \{I\} \}$.
We lift $\decode \dr$ and $\encode \dr$ to sets of interpretations in the standard manner.
\begin{example}
We consider the domain restriction $\dr$ such that $\dr(x)	 =(A_x,\true)$ from Figure~\ref{fig:sfa}~(a) and $\dr(y) = (A_y,\true)$ from Figure~\ref{fig:sfa}~(b). We assume that the set of integer variables $Z=\{v_z\}$ and $V_{\dr(x)} =\{v_0, v_1,v_2,v_3\}$. For the interpretation $I_e =(v_z,v_0,v_1,v_2,v_3,\#v_0,\#v_1,\#v_2,\#v_3) \rightarrow (3,\semof{a}, \semof{b}, \semof{c},\semof{\epsilon},3,3,2,2)$, we have $\decode \dr (I_e) = \{(z,x,y) \rightarrow(3, ``ababab",``cc") \}$. In the other direction, for $I=(z,x,y) \rightarrow(3, ``ababab",``cc")$, we have $\encode \dr (I) =\{
(v_z,v_0,v_1,v_2,v_3,\#v_0,\#v_1,\#v_2,\#v_3) \rightarrow (3,\semof{a}, \semof{b}, \semof{c},\semof{\epsilon},3,3,2,2), (v_z,v_0,v_1,v_2,v_3,\#v_0,\#v_1,\#v_2,\#v_3) \rightarrow (3,\semof{a}, \semof{b}, \semof{c},\semof{c},3,3,1,1), \ldots\}$

\end{example}

\begin{theorem}\label{theorem:correct}
$\decode \dr (\semof {\underf \dr {\insc}}) = \semof{\insc}^\dr$ 
\end{theorem}

@@

Formally, the asynchronous product automaton is a tuple $\syncof \pa {\pa'} = (Q\times Q', \syncT, \cvarone_\epsilon \times \cvartwo_\epsilon, (q_i,q_i'),(q_f,q_f'))$, where the transition relation $\syncT$ is the minimal set satisfying the following:
%
\begin{itemize}
\item If $(q_1,v,q_2) \in T$ and $(q_1',v',q_2') \in T'$, then we have $((q_1,q_1'),(v,v'),(q_2,q_2'))\in \syncT$.

\item If $(q_1,v,q_2) \in T$, then for all states $q'\in Q'$, we have $((q_1,q'),(v,\epsilon),(q_2,q'))\in \syncT$.

\item If $(q_1',v',q_2') \in T'$, then for all states $q\in Q$, we have $((q,q_1'),(\epsilon,v'),(q,q_2'))\in \syncT$.
\end{itemize}	
%
The synchronization formula $\syncfof{\pa} {\pa'}$ is extracted from $\syncof \pa {\pa'}$ as follows.
Its first part is the Parikh formula $\parikhfof {\syncof \pa {\pa'}}$ of the product, which encodes all accepting runs of $\syncof \pa {\pa'}$.
%
The second part is a constraint that extracts from a run of $\syncof \pa {\pa'}$ the corresponding runs of $\pa$ and of $\pa'$:
$$ \Psi_{\#} \defeq 
\left(\bigwedge_{v\in\cvarone}\#v = \sum_{x' \in \cvartwo_\epsilon} \#{(v,x')}\right)
\land
\left(\bigwedge_{v'\in\cvartwo} \#v' = \sum_{x \in \cvarone_\epsilon} \#{(x,v')}\right)
$$
Notice that $x,x'$ are either variables or $\epsilon$.
Finally, the third part forces the interpretations of the parametric words accepted by $\pa$ and $\pa'$ to be the same:
$$ \Psi_= \defeq
\bigwedge_{x\in\cvarone_\epsilon,x'\in\cvartwo_\epsilon} \#(x,x')>0 \rightarrow (x=x')
$$
The synchronization formula is then the conjunction 
$$
\syncfof \pa {\pa'} \defeq \parikhfof{\syncof \pa {\pa'}} \land \Psi_{\#} \land \Psi_= \land \paf \land \paf'
$$
